\documentclass{article}
\usepackage{amsmath,amssymb,braket,url,hyperref,booktabs,graphicx}
\usepackage[style=nature]{biblatex}
\addbibresource{12_ref.bib}
\newcommand{\bfit}[1]{\textit{\textbf{#1}}}
\begin{document}
\medskip
\noindent
\textbf{\Large Chapter 12\\
Electron Spin Qubit in Semicoductor-1-Qubit and 2-Qubit Gates}\\\\\\
\medskip
\textbf{\large 12.1 Introduction}\\
In Chap.11, we showed how to implement a qubit using electron spin on a silicon
substrate. We also demonetrated how to initialize and measure a qubit. In this chapter,
we will sudy how to perform a universal 1-qubit gate and a 2-qubit entnaglement gate to fulfill
the last two DiVincenzo's criteria (Sect. 1.3). In Chap.10, we showed
that by applying a vertical DC magnetic field and a rotating  horizontal magnetic field and
then \textit{working in the rotating frame}, we would be able to rotate any state on Bloch sphere
about any vector. This allows us to build a universal 1-qubit gate (Section 27.4 in \cite{WongHuiYong})
However, in the literature, many silicon qubits are still implemented with the setup in Chap. 9 which means
that the qubit is places in a vertical DC magnetic field and a perturbating and linearly oscillating horizontal
magnetic field. This is what we will use in this chapter. We will first summarize an experimental paper on how
it implements 1-qubit gate. Then we will discuss the implementation of a 2-qubit entnaglement gate with an example.
\\\\\\
\bfit{\large 12.1.1 Learning Outcomes}
\\\\
Be able to describe how a 1-qubit gate can be implemented for silicon spin qubits; 
understant how a CNOT-gate can be impelmented by using a native entanglement gate of
silicon spin qubits and other 1-qubit gates.
\\\\
\bfit{\large 12.1.2 Teaching Videos}\\\\
$\bullet$ Search for Ch12 in this playlist\\

- \url{https//tinyurl.com/3yhze3jn}\\\\
$\bullet$ Other Videos\\

- \url{https://youtu.be/0JVw4xICV10}

- \url{https//youtu.be/_CpQ-Uy0Kgo}
\\\\\\
\textbf{\large 12.2 1-Qubit Gate Implementation}
\\\\
We will use an example in \cite{veldhorst2014addressable} (with some variation)
to demonstrate how to implement electron spin qubit in silicon and the 1-qubit gate.
The setup is illustrated in Fig. 12.1.

Firstly, the Hilbert space is created by applying a DC magnetic field pointing to the right.
This direction is named the $-\hat{z}$ direction. Therefore, \textit{spin-up means that the spin is 
pointing to the left, and spin-down means that the spin is pointing to the right}. This is nothing
special becuase the name of direction is completely a human definition.
\printbibliography
\end{document}
